Design and Implementation of Probabilistic Programming Language Anglican

### 2016-11-27 paper/paper.tex
update

Anglican can be mostly implemented as a regular programming
language, except for the handling of these checkpoints.
Depending on the \textit{inference algorithm}, \texttt{sample}
and \texttt{observe} may result in implicit input/output
operations and control changes. For example, \texttt{observe} in
particle filtering inference algorithms \cite{WVM14} is a non-deterministic
control statement at which a particle (corresponding to
a thread executing a program)
can be either replicated or terminated. Similarly, in 
Metropolis-Hastings \cite{WSG11},
\texttt{sample} is both an input
statement which `reads' values from a random source,
and a non-deterministic control statement
(with delayed effect), eventually affecting acceptance or
rejection of a sample.

@@

Typically, an inference
algorithm has its own implementations of
\texttt{checkpoint} for \texttt{sample}, \texttt{observe}, or
both, as well as invoke \texttt{exec} from an elaborated
conditional control flow. LMH (\texttt{anglican.lmh}) and SMC
(\texttt{anglican.smc}) are examples of inference algorithms
where either \texttt{observe} (SMC) or \texttt{sample} (LMH)
handler is overridden. In addition, SMC runs multiple particles
(threads executing a program) simultaneously, while LMH re-runs
programs from an intermediate continuation rather than from the
beginning.

@@

Many inference algorithms memoize and reuse earlier computations
at checkpoints: variants of Metropolis-Hastings reuse previously
drawn values~\cite{WSG11}, as well as additional information
used for adaptive sampling~\cite{TMP+15} at \texttt{sample}
checkpoints, Asynchronous SMC (Particle Cascade) computes
average particle weights at \texttt{observe}
checkpoints~\cite{PWD+14}. Implementations of black-box
variational inference~\cite{WW13,MPT+16} associate with random
variables the learned parameters of variational posterior
distribution.

>>>>

Anglican can be mostly implemented as a regular programming
language, except for the handling of these checkpoints.
Depending on the \textit{inference algorithm}, \texttt{sample}
and \texttt{observe} may result in implicit input/output
operations and control changes. For example, \texttt{observe} in
particle filtering inference algorithms \cite{WVM14} is a non-deterministic
control statement at which a particle (corresponding to
a user-level thread executing a program)
can be either replicated or terminated. Similarly, in 
Metropolis-Hastings \cite{WSG11},
\texttt{sample} is both an input
statement which `reads' values from a random source,
and a non-deterministic control statement
(with delayed effect), eventually affecting acceptance or
rejection of a sample.

@@

Typically, an inference
algorithm has its own implementations of
\texttt{checkpoint} for \texttt{sample}, \texttt{observe}, or
both, as well as invoke \texttt{exec} from an elaborated
conditional control flow. LMH (\texttt{anglican.lmh}) and SMC
(\texttt{anglican.smc}) are examples of inference algorithms
where either \texttt{observe} (SMC) or \texttt{sample} (LMH)
handler is overridden. In addition, SMC runs multiple particles
(corresponding to user-level threads) simultaneously, while LMH re-runs
programs from an intermediate continuation rather than from the
beginning.

@@

Many inference algorithms memoize and reuse earlier computations
at checkpoints. Variants of Metropolis-Hastings reuse previously
drawn values~\cite{WSG11}, as well as additional information
used for adaptive sampling~\cite{TMP+15} at \texttt{sample}
checkpoints. Asynchronous SMC (Particle Cascade) computes
average particle weights at \texttt{observe}
checkpoints~\cite{PWD+14}. Implementations of black-box
variational inference~\cite{WW13,MPT+16} associate with random
variables the learned parameters of variational posterior
distribution.
